Examples of zero-inflated negative binomial regression

Example 1. School administrators study the attendance behavior of high school juniors at two schools. Predictors of the number of days of absence include gender of the student and standardized test scores in math and language arts.

Example 2. The state wildlife biologists want to model how many fish are being caught by fishermen at a state park. Visitors are asked how long they stayed, how many people were in the group, were there children in the group and how many fish were caught. Some visitors do not fish, but there is no data on whether a person fished or not. Some visitors who did fish did not catch any fish so there are excess zeros in the data because of the people that did not fish. 

Description of the Data

Let's pursue Example 2 from above.
 
We have data on 250 groups that went to a park. Each group was questioned about how many fish they caught (count), how many children were in the group (child), how many people were in the group (persons), and whether or not they brought a camper to the park (camper).

In addition to predicting the number of fish caught, there is interest in predicting the existence of excess zeros, i.e., the probability that a group caught zero fish. We will use the variables child, persons, and camper in our model. Let's look at the data.



zinb <- read.csv("http://www.ats.ucla.edu/stat/data/fish.csv")
zinb <- within(zinb, {
  nofish <- factor(nofish)
  livebait <- factor(livebait)
  camper <- factor(camper)
})

summary(zinb)
 
##  nofish  livebait camper     persons         child             xb        
##  0:176   0: 34    0:103   Min.   :1.00   Min.   :0.000   Min.   :-3.275  
##  1: 74   1:216    1:147   1st Qu.:2.00   1st Qu.:0.000   1st Qu.: 0.008  
##                           Median :2.00   Median :0.000   Median : 0.955  
##                           Mean   :2.53   Mean   :0.684   Mean   : 0.974  
##                           3rd Qu.:4.00   3rd Qu.:1.000   3rd Qu.: 1.964  
##                           Max.   :4.00   Max.   :3.000   Max.   : 5.353  
##        zg             count      
##  Min.   :-5.626   Min.   :  0.0  
##  1st Qu.:-1.253   1st Qu.:  0.0  
##  Median : 0.605   Median :  0.0  
##  Mean   : 0.252   Mean   :  3.3  
##  3rd Qu.: 1.993   3rd Qu.:  2.0  
##  Max.   : 4.263   Max.   :149.0
 
## histogram with x axis in log10 scale
ggplot(zinb, aes(count, fill = camper)) +
  geom_histogram() +
  scale_x_log10() +
  facet_grid(camper ~ ., margins=TRUE, scales="free_y")
 
## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
## stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this.
 
